---
title: "Puffin colour banding 2016-2021"
author: "Anna"
date: "`r format(Sys.time(), '%d %B, %Y %H:%M:%S')`"
output: html_document
editor_options:
  chunk_output_type: console
---

Notes about the data set: These models are now just running the truncated colour-banding data, from 2016-2021, given that there were so many earlier years with no recapture effort. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```
```{r, include=FALSE}

#Load project settings
library(here)
source(here::here("project_settings.R"), echo=T)

```
###############################################################
### 1- Import data and format for RMARK
###############################################################
```{r}

#bring in the inp file from Erpur, but we'll want to only keep digits 9-14 from the ch string:
puff <- convert.inp(here(CMR.dir, "colour banding/Puffin Colour SH 2021.inp"))
puff.recent <- puff
puff.recent$ch <- substr(puff$ch, 9, 14)
#and then need to delete any rows with all 0s
puff.recent <- puff.recent %>% filter(!ch =="000000")
nrow(puff.recent)
#So there are now 182 banded birds in the data set from 2016-2021

puff.recent.processed <- process.data(puff.recent, model = 'CJS', begin.time=2016)
puff.recent.ddl <- make.design.data(puff.recent.processed)

# add tsm structure
puff.recent.ddl <- add.design.data(puff.recent.processed, puff.recent.ddl,
                            parameter="Phi", type="age", bins=c(0,1, 17),name="tsm",
                            right = F, replace = T)

summary_ch(puff.recent.processed, marray = T)


```
###############################################################
### 2- Set up parameter variation structure and run general model
###############################################################
```{r}
#parameter variation structure  (include tsm from the start)
Phi.dot <- list(formula = ~1)
Phi.t <- list(formula = ~time)
Phi.tsm <- list(formula =~ tsm)
Phi.tsmplust <- list(formula = ~tsm + time)
p.dot <- list(formula = ~1)
p.t <- list(formula = ~time)
```
##############################################################
### 3- Goodness of Fit testing
##############################################################
```{r}
#RELEASE/UCARE tests
# export capture history as inp and import again to use with R2ucare
export.chdata(puff.recent.processed, filename="puffin colourband 2016-21.ch" ,replace=TRUE)

puff.recent.inp = read_inp("puffin colourband 2016-21.ch")

puff.recent.hist = puff.recent.inp$encounter_histories
puff.recent.freq = puff.recent.inp$sample_size


overall_CJS(puff.recent.inp$encounter_histories, puff.recent.inp$sample_size) #significant
#Not significant
test3sr(puff.recent.inp$encounter_histories, puff.recent.inp$sample_size)
test3sm(puff.recent.inp$encounter_histories, puff.recent.inp$sample_size)
test2ct(puff.recent.inp$encounter_histories, puff.recent.inp$sample_size)
test2cl(puff.recent.inp$encounter_histories, puff.recent.inp$sample_size)
#all not significant

##Bootstrap GOF 
#first we need to build a general model 
global<-mark(puff.recent.processed, puff.recent.ddl, model="CJS", model.parameters=list(Phi=Phi.tsmplust, p=p.t))
cat("global model: ",global$model.name)

#And now we should be able to just call Greg's bootstrap function to run on this model
source(file.path(R.dir, "CMR/CMR utilities/bootstrap_deviance.R"))
bootstrap.deviance(global, 10, tsm=TRUE)
#working fine here (not sure why it wouldn't work for 2008-21 dataset?)
#gives c-hat lower than 1 (0.9666), and CJS was also non significant, so not making any adjustment for overdispersion

```
##############################################################
### 4- Run full model set
##############################################################
```{r}
#create whole model set
model.list=create.model.list("CJS") # models to test
cat("number of models in candidate set:", nrow(model.list))  #  8 models total

all.models=mark.wrapper(model.list,data=puff.recent.processed,ddl=puff.recent.ddl,threads=2,output=F,silent=T)
options(scipen = 999) # disable scientific notation for easier reading
results <- all.models$model.table

kbl(results) %>% kable_styling(bootstrap_options = c("striped", "hover"), fixed_thead = T)  %>% scroll_box(width = "100%", height = "500px")

```

##############################################################
### 5- Explore specific estimates from top models
#############################################################
Top-ranked model is Phi(tsm+t)p(.) but many models very closely ranked, and time-variant-Phi models include mostly boundary estimates (i.e. not estimable)

```{r}
top.model <- select_model_results(model_num = 4)
top.model$model.name
PIMS(top.model, parameter = "Phi")
round(top.model$results$real[,1:4],4)
#lots of boundary estimates - not really much of value in time-variant

#so look at time-invariant survival, with or without tsm (transient) effect:

Phidotpdot.model <- select_model_results(model_num=1)
Phidotpdot.model$model.name
PIMS(Phidotpdot.model, parameter = "Phi")
round(Phidotpdot.model$results$real[,1:4],4)
# mean survival 0.971 without transient effect

Phitsmpdot.model <- select_model_results(model_num=7)
Phitsmpdot.model$model.name
PIMS(Phitsmpdot.model, parameter = "Phi")
round(Phitsmpdot.model$results$real[,1:4],4)
#non-transient survival estimated at 1 (boundary, i.e. not estimated properly)

```
### MAIN RESULT: The best estimate of mean adult puffin survival from the truncated colour banding data set, using only 2016-2021 data, is 0.971. Note, however, that many parameters were poorly estimated in this short data set, and that a tsm (transient) model did not produce a good estimate of survival.

```{r, include=FALSE}
### IMPORTANT LAST STEP: get rid of junk MARK files!
force.cleanup(recursive = TRUE)
```
