---
title: "Puffin colour banding 2008-2021"
author: "Anna"
date: "`r format(Sys.time(), '%d %B, %Y %H:%M:%S')`"
output: html_document
editor_options:
  chunk_output_type: console
---

### Notes about the data set: These are resightings of colour-banded breeding birds at Storhofdi. There were no resightings in 2010, 2011, 2012, 2014 and 2015 (but note that there were still birds banded in 2014 and 2015). [For comparison, the Helgason 2012 thesis analysis was for recaptures, not resights, from 1959-2007; its best model was a time-invariant-tsm model, with mean survival estimated at 0.873 for non-transients]
--------------------------------------------------------------------------------


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```
```{r, include=FALSE}

#Load project settings
library(here)
source(here::here("project_settings.R"), echo=T)
```
###############################################################
### 1- Import data and format for RMARK
###############################################################
```{r}
#bring in the inp file that Erpur sent
puff <- convert.inp(here(CMR.dir, "colour banding/Puffin Colour SH 2021.inp"))

#number of individual encounter histories
nrow(puff)

puff.processed <- process.data(puff, model = 'CJS', begin.time=2008)
puff.ddl <- make.design.data(puff.processed)


# add tsm structure
puff.ddl <- add.design.data(puff.processed, puff.ddl,
                            parameter="Phi", type="age", bins=c(0,1, 17),name="tsm",
                            right = F, replace = T)

summary_ch(puff.processed, marray = T)
##note that not only was there no recapture at all in 2010/11/12/14/15 but also very low in 2009, 2013
#so maybe i'll end up wanting to try a truncated analysis, e.g. starting w releases in 2014-onward?
```
 
###############################################################
### 2- Set up parameter variation structure and run general model
###############################################################

```{r}

#parameter variation structure  (include tsm from the start)
Phi.dot <- list(formula = ~1)
Phi.t <- list(formula = ~time)
Phi.tsm <- list(formula =~ tsm)
Phi.tsmplust <- list(formula = ~tsm + time)

#fix p=0 for years where there was no recap effort (2010,11,12,14,15)
p.dot=list(formula=~1, fixed=list(time=c(2010, 2011, 2012, 2014, 2015), value=0)) #no variation over time [except still 0 in no-effort years] 
p.t <- list(formula=~ time, fixed=list(time=c(2010, 2011, 2012, 2014, 2015), value=0)) #varies among years but still fixed to 0 in all those years

```

##############################################################
### 3- Goodness of Fit testing
##############################################################

```{r}

#RELEASE/UCARE tests
puff.inp <- read_inp(here(CMR.dir, "colour banding/Puffin Colour SH 2021.inp"))


overall_CJS(puff.inp$encounter_histories, puff.inp$sample_size) #significant
#Not significant
test3sr(puff.inp$encounter_histories, puff.inp$sample_size)
test3sm(puff.inp$encounter_histories, puff.inp$sample_size)
test2ct(puff.inp$encounter_histories, puff.inp$sample_size)
test2cl(puff.inp$encounter_histories, puff.inp$sample_size)
#all non significant

##Bootstrap GOF 
#first we need to build a general model 
global<-mark(puff.processed, puff.ddl, model="CJS", model.parameters=list(Phi=Phi.tsmplust, p=p.t))
cat("global model: ",global$model.name)

#And now we should be able to just call Greg's bootstrap function to run on this model
## GOt an error message, didn't work
#source(file.path(R.dir, "CMR/CMR utilities/bootstrap_deviance.R"))
#bootstrap.deviance(global, 10, tsm=TRUE)

#Run bootstrap in MARK instead: 100 reps for model Phi(tsm+t)p(t) with p=0 for 2010-11-12-14-15
#observed model deviance = 98.944553, mean simulated deviance = 84.253
c.hat <- 98.94453/84.253
c.hat

```
##############################################################
### 4- Run full model set
##############################################################
```{r}
#create whole model set
model.list=create.model.list("CJS") # models to test
cat("number of models in candidate set:", nrow(model.list))  #  8 models total

all.models=mark.wrapper(model.list,data=puff.processed,ddl=puff.ddl,threads=2,output=F,silent=T)
options(scipen = 999) # disable scientific notation for easier reading

#adjust the model results table with the bootstrap c-hat and modify the table
all.models <- adjust.chat(c.hat,all.models)
results <- all.models$model.table

kbl(results) %>% kable_styling(bootstrap_options = c("striped", "hover"), fixed_thead = T)  %>% scroll_box(width = "100%", height = "500px")

```
##############################################################
### 5- Explore specific estimates from top models
#############################################################
Top-ranked model is Phi(tsm)p(t), i.e. time-invariant tsm model with time-variant recapture
```{r}
#top model is Phi(tsm)p(t) , and Phi(dot)p(t) is second, i.e. no time var in Phi but lots in p

top.model <- select_model_results(model_num = 7)
top.model$model.name
PIMS(top.model, parameter = "Phi")
round(top.model$results$real[,1:4],4)
## gives non-transient survival of 0.956, good!

#without the transience effect, get survival estimated =1? (boundary)
second.model <- select_model_results(model_num=5)
second.model$model.name
PIMS(second.model, parameter="Phi")
round(second.model$results$real[,1:4,4])
## Doesn't work! (boundary estimates at 0 or 1)

```
### MAIN RESULT: The best estimate of adult puffin survival from the 2008-21 colour-banding data is 0.956 (SE 0.32; non-transient individuals), with high inter-annual variation in recapture probability (ranging between 0.05-0.27; fixed to 0 in years 2010-12, 2014, 2015). 

```{r, include=FALSE}
### IMPORTANT LAST STEP: get rid of junk MARK files!
force.cleanup(recursive = TRUE)
```